---
title: "PM 2.5 in Germany"
output: html_notebook
---

```{r}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(plotly)
library(grid) #for adjust grid scale
```

```{r}
#import data from url or file

# Chiang Mai - http://berkeleyearth.lbl.gov/air-quality/maps/cities/Thailand/Chiang_Mai/Chiang_Mai.txt
# Bangkok - http://berkeleyearth.lbl.gov/air-quality/maps/cities/Thailand/Bangkok/Bangkok.txt

url <- 'http://berkeleyearth.lbl.gov/air-quality/maps/cities/Thailand/Chiang_Mai/Chiang_Mai.txt'
df <- read_tsv(url,skip=10, col_names = FALSE)
df
```

```{r}
colnames(df) <- c('Year','Month','Day','Hour','PM2_5','X6','X7')
df
```

```{r}
df <- df %>%
  select(Year:PM2_5)
df
```

```{r}
df <- df %>% 
  mutate(date_time = ISOdate(Year,Month,Day,Hour),
         local_dt = date_time+hours(7),
         local_hr = hour(local_dt))
df
```
 ------ Summarized Data (Reserved) ------

```{r}
df %>% 
  mutate(year = year(local_dt),
         month = month(local_dt),
         date = date(local_dt)) %>%
  group_by(date) %>%
  summarize(avg_pm = mean(PM2_5))
```
 ------ Visualized Data ------
 
 ## entire data
```{r}
df %>% 
  mutate(year = year(local_dt),
         month = month(local_dt),
         date = date(local_dt)) %>%
  group_by(date) %>%
  summarize(avg_pm = mean(PM2_5)) %>%
  ggplot(aes(date, y = avg_pm , color = as.factor(year(date))))+geom_line()+geom_smooth()
```

## by date for each month
```{r}
df %>% 
  mutate(year = year(local_dt),
         month = month(local_dt),
         date = date(local_dt)) %>%
  group_by(year,month,date) %>%
  summarize(avg_pm = mean(PM2_5)) %>%
  filter(year == 2019) %>%
  ggplot(aes(date, y = avg_pm , color = as.factor(month)))+geom_line()
```

## By month for each year
```{r}
df %>% 
  mutate(year = year(local_dt),
         month = month(local_dt),
         date = date(local_dt)) %>%
  group_by(year,month) %>%
  summarize(avg_pm = mean(PM2_5)) %>%
  ggplot(aes(month, y = avg_pm , color = as.factor(year)))+ geom_smooth(se=FALSE) +scale_x_discrete(limits =month.abb)  
```
 
 ## create heat map
```{r}
df %>% 
  mutate(date = date(local_dt)) %>%
  group_by(date) %>%
  summarize(avg_pm = mean(PM2_5)) %>%
  mutate(day = day(date),
         month = month(date),
         year = year(date)) %>%
  ggplot(aes(day, y=month, fill=avg_pm)) + geom_tile()
```
 
 ## add factor to map, make it display by day
 
```{r}
df %>% 
  mutate(date = date(local_dt)) %>%
  group_by(date) %>%
  summarize(avg_pm = mean(PM2_5)) %>%
  mutate(day = as.factor(day(date)),
         month = as.factor(month(date)),
         year = as.factor(year(date))) %>%
  ggplot(aes(day, y=month, fill=avg_pm)) + geom_tile()
```
 
 ## add border and set a color
```{r}
df %>% 
  mutate(date = date(local_dt)) %>%
  group_by(date) %>%
  summarize(avg_pm = mean(PM2_5)) %>%
  mutate(day = as.factor(day(date)),
         month = as.factor(month(date)),
         year = as.factor(year(date))) %>%
  ggplot(aes(day, y=month, fill=avg_pm)) +
  geom_tile(color="white") +
  scale_fill_gradient(low = "green" , high = 'red')
```
 
## adjust cell scale
```{r, fig.hight=2, fig.width=8} 
df %>% 
  mutate(date = date(local_dt)) %>%
  group_by(date) %>%
  summarize(avg_pm = mean(PM2_5)) %>%
  mutate(day = as.factor(day(date)),
         month = as.factor(month(date)),
         year = as.factor(year(date))) %>%
  ggplot(aes(day, y=month, fill=avg_pm)) +
  geom_tile(color="white") +
  scale_fill_gradient(low = "green" , high = 'red') +
  coord_equal()
```
## set interval factor for avg pm 2.5

```{r}
df %>% 
  mutate(date = date(local_dt)) %>%
  group_by(date) %>%
  summarize(avg_pm = mean(PM2_5)) %>%
  mutate(day = as.factor(day(date)),
         month = as.factor(month(date)),
         year = as.factor(year(date))) %>%
  mutate(pm_lv = cut(avg_pm, ## create new column to classify level of smog
         breaks = c(0,12,35.4,55.4,150.4,250.4,350.4),
         labels = c('good','moderate','unheathy for sensitive groups'
                    ,'unheathy','very unheathy','hazardous'))) %>%
  ggplot(aes(day, y=month, fill=pm_lv)) + ## replace avg_pm with pm_level
  geom_tile(color="white") +
  scale_fill_manual(values = c('green','yellow','orange','red','magenta', 'violet'))+
  coord_equal()
```

## separate by year
```{r,fig.hight=12, fig.width=16}
df %>% 
  filter(year(local_dt) > 2015) %>%
  mutate(date = date(local_dt)) %>%
  group_by(date) %>%
  summarize(avg_pm = mean(PM2_5)) %>%
  mutate(day = as.factor(day(date)),
         month = as.factor(month(date)),
         year = as.factor(year(date))) %>%
  mutate(pm_lv = cut(avg_pm, 
         breaks = c(0,12,35.4,55.4,150.4,250.4,350.4),
         labels = c('good','moderate','unheathy for sensitive groups'
                    ,'unheathy','very unheathy','hazardous'))) %>%
  ggplot(aes(day, y=month, fill=pm_lv)) + 
  geom_tile(color="white") +
  facet_grid(year~.) + ## separate by year
  scale_fill_manual(values = c('green','yellow','orange','red','magenta', 'violet'))+
  coord_equal()

```
## reverse month -> make less value on top
```{r,fig.hight=12, fig.width=16}
df %>% 
  filter(year(local_dt) > 2015) %>%
  mutate(date = date(local_dt)) %>%
  group_by(date) %>%
  summarize(avg_pm = mean(PM2_5)) %>%
  mutate(day = as.factor(day(date)),
         month = fct_rev(as_factor(month(date))),  ## remove as.factor and add fct_rev() to month and make it factor using as_factor
         year = as.factor(year(date))) %>%
  mutate(pm_lv = cut(avg_pm, 
         breaks = c(0,12,35.4,55.4,150.4,250.4,350.4),
         labels = c('good','moderate','unheathy for sensitive groups'
                    ,'unheathy','very unheathy','hazardous'))) %>%
  ggplot(aes(day, y=month, fill=pm_lv)) + 
  xlab("Day") + ylab("Month") +
  geom_tile(color="white") +
  facet_grid(year~.) + 
  scale_fill_manual(values = c('green','yellow','orange','red','magenta', 'violet'))+
  coord_equal()
```
## add tooltip

```{r}
tt <- df %>%  ## put ggplot in to tt (tooltip)
  filter(year(local_dt) > 2017) %>%
  mutate(date = date(local_dt)) %>%
  group_by(date) %>%
  summarize(avg_pm = mean(PM2_5)) %>%
  mutate(day = as.factor(day(date)),
         month = fct_rev(as_factor(month(date))),  
         year = as.factor(year(date))) %>%
  mutate(pm_lv = cut(avg_pm, 
         breaks = c(0,12,35.4,55.4,150.4,250.4,350.4),
         labels = c('good','moderate','unheathy for sensitive groups'
                    ,'unheathy','very unheathy','hazardous'))) %>%
  ggplot(aes(day, y=month, fill=pm_lv)) + 
  xlab("Day") + ylab("Month") +
  geom_tile(color="white") +
  facet_grid(year~.) + 
  scale_fill_manual(values = c('green','yellow','orange','red','magenta', 'violet'))+
  coord_equal()

ggplotly(tt) # display ggplot with tooltip
```


## figure out scale error by increase number of column and set y's scale to free
```{r}
tt<-df %>% 
  filter(Year > 2015) %>%
  mutate(date = date(local_dt)) %>%
  group_by(date) %>%
  summarize(avg_pm = mean(PM2_5)) %>%
  mutate(day = as.factor(day(date)),
         month = fct_rev(as_factor(month(date))),
         year = as.factor(year(date)),
         pm_lv = cut(avg_pm, 
           breaks = c(0,12,35.4,55.4,150.4,250.4,350.4),
           labels = c('good','moderate','unheathy for sensitive groups'
                      ,'unheathy','very unheathy','hazardous'))
         ) %>%
  ggplot(aes(day, y=month, fill=pm_lv)) + 
  xlab("Day") + ylab("Month") +
  geom_tile(color="white") +
  facet_wrap(year~., scales = "free", ncol=2) + 
  scale_fill_manual(values = c('green','yellow','orange','red','magenta', 'violet'))

  ggplotly(tt)
```

# like above but force square coord_equal()
```{r}
tt<-df %>% 
  filter(Year > 2018) %>%
  mutate(date = date(local_dt)) %>%
  group_by(date) %>%
  summarize(avg_pm = mean(PM2_5)) %>%
  mutate(day = as.factor(day(date)),
         month = fct_rev(as_factor(month(date))),
         year = as.factor(year(date)),
         pm_lv = cut(avg_pm, 
           breaks = c(0,12,35.4,55.4,150.4,250.4,350.4),
           labels = c('good','moderate','unheathy for sensitive groups'
                      ,'unheathy','very unheathy','hazardous'))
         ) %>%
  ggplot(aes(day, y=month, fill=pm_lv)) + 
  xlab("Day") + ylab("Month") +
  geom_tile(color="white") +
  facet_wrap(year~., scales = "free", ncol=1) + 
  scale_fill_manual(values = c('green','yellow','orange','red','magenta', 'violet'))+
  coord_equal()

  ggplotly(tt)
```